\documentclass{sbthesis} % source: Dr. Keith Schubert
%\usepackage[pdftex]{graphicx}
\usepackage{graphicx}
\usepackage{verbatim}
\usepackage{textcomp}
\usepackage{setspace}
\usepackage{float}
\usepackage{url}
\usepackage{regexpatch}% http://ctan.org/pkg/regexpatch
\usepackage[titletoc]{APPENDIX}
\usepackage{listings}
\usepackage{caption}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% NOTE: I didn't have to do this.  Maybe it's because I'm using Miktex.  -- David Turner
%
% Note: bibtex omits page number from first page of bibliography, 
% which is not acceptable under CSUSB rules.  The following
% command can be used to fix the problem.  For this to work,
% you need to insert \FixBib after the first line in the bbl file 
% that bibtex generates.  Do this each time the bibliography
% changes.
%
% Source: http://www.cs.ubc.ca/~murphyk/Teaching/latex_tips.txt
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand{\FixBib}{%    PUT \FixBib in file.bbl after first line
    \setlength{\parsep}{\parskip}%
    \setlength{\itemsep}{0cm}%
    \setlength{\topsep}{\parskip}%
    \setlength{\parskip}{0cm}%
    \setlength{\partopsep}{0cm}%
    \setlength{\listparindent}{\parindent}%
    \setlength{\labelwidth}{10pt}%
    \setlength{\labelsep}{0pt}%
    \setlength{\leftskip}{0pt}%
    \setlength{\leftmargin}{0pt}%
}

\title{GRADEBADGE:}
\TitleLineTwo{Development of a Cloud-Based Reward Application}
\author{Erwin Toni Soekianto}
\Department
\EmailAddress{soekiane@coyote.csusb.edu}
\Advisor{David Turner}
\Committee{Richard J. Botting}{Arturo I. Concepcion}
\CSUSBDate{June 2013}


\AbstractText{
The purpose of this project is to investigate the use of cloud-based services to deliver cutting-edge applications. For this purpose, a prototype of a reward application using badges, called GradeBadge, was developed to illustrate and explore this emerging paradigm. The cloud services utilized by this application include Heroku (for the application server), MongoLab (for the database), and Facebook (for authentication and social network integration).  

The application server is written in Javascript and runs inside a Nodejs execution environment. The application is accessed through a Web browser running on either desktop or mobile computers. On the client side, this application makes use of numerous Web technologies, including HTML5, CSS, Bootstrap, and Jquery. The project also made use of the Git version control system to manage source code and deployment of the application server to Heroku. The source code repository was stored remotely through the cloud-based service called GitHub. 

The purpose of the GradeBadge application is to help organizations interact with and motivate their members in a fun way. It keeps the members engaged by giving badges as rewards for their efforts or achievements. In order to facilitate adoption among users, GradeBadge is integrated with the social networking site Facebook.

As the result of this implementation in GradeBadge using cloud computing, we don't need to spend time on system administration to manage the servers. We also don't need to purchase and maintain the hardware. Deployment and re-deployment are done easily from the local command line. For this reason, our implementation approach is suitable with student projects or small startup companies.

Cloud computing provides significant cost savings to developers when building applications that can be scaled up or down almost instantly to accomodate rapidly changing demand.
}

\AcknowledgementText{
I would like to thank all the people with whom I have worked while pursuing my master's degree at California State University, San Bernardino (CSUSB). Studying in the School of Computer Science and Engineering at CSUSB has been a great learning experience.  I would like to thank the faculty of the School of Computer Science and Engineering who supported this project by serving on my committee: Dr. David Turner, Dr. Arturo Concepcion and Dr. Richard Botting.
}

\begin{document}
\Project  % defined in sbthesis.cls 

\input{introduction}
\input{requirements}
\input{architecture}
\input{system_design}
\input{database_design}
\input{userinterface_design}
\input{conclusion}
%\input{appendixA}
%\input{appendixB}
%\input{appendixC}
%\input{appendixD}

%\appendix  not used, from Dr. Schubert's template
\appendix
%\input{appendixA}
%\input{appendixB}
%\input{appendixC}
%\input{appendixD}

\nocite{*}
\urlstyle{same}
\raggedright
\sloppy
%\renewcommand{\normalsize}{\fontsize{12pt}{12pt}\selectfont} 
\normalsize
\Bibliography{main}

\end{document}
